---
title: "Case Study 8 & 9"
output: html_document
---

```{r setup, include=FALSE}
#Bring in Libraries

library(plyr)     
library(tidyr)   
library(dplyr)

```

## R Markdown

Bring in Data

```{r cars}

Beers_df=read.csv(file.choose(),header=TRUE)
Breweries_df=read.csv(file.choose(),header=TRUE)

```


Question 1: How many breweries are present in each state?

```{r pressure, echo=FALSE}
Breweries_State <-aggregate(cbind(count = Name) ~ State,data = Breweries_df,FUN=function(x){NROW(x)})

head(Breweries_State,2)
```





Question 2   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)

```{r pressure, echo=FALSE}

#change the brew ID's header  from the Brewery_Table 
Breweries_df <- Breweries_df %>% rename(Brewery_id=Brew_ID)

#change the name column to Brewery_ID.Name to avoid duplicates columns
Breweries_df <- Breweries_df %>% rename(Brewery_id.Name=Name)

#Right join to the brewery data
CombinedData<-join(Beers_df,Breweries_df,by ="Brewery_id",type = "right")

head(CombinedData,2)
```

## Including Plots




Question 3: Address the missing values in each column.

```{r cars}

CombinedData[CombinedData=='']<-NA

colSums(is.na(CombinedData))

```


Question 4: Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.


```{r cars}

#changed #NA to 0 to get actual medians
CombinedData[is.na(CombinedData)]<-0

#created the dataframes with medians for each State
Medians<-CombinedData%>%group_by(State)%>%summarise(median.ABV=median(ABV),median.IBU=median(IBU))

head(Medians,6)

```



Question 5: Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

the state with the highest ABV = Colorado
the state with the highest IBU = Oregon

```{r cars}

Maximum<-CombinedData%>%group_by(State)%>%summarise(Max.ABV=max(ABV),Max.IBU=max(IBU))

head(Maximum,5)
```


Question 6: Comment on the summary statistics and distribution of the ABV variable.

- ABV is normally in the range of 0.001 to 0.128. 
- But most of the beers in the country have a ABV of 0.059

- Based on the histogram and qqplot we can see that the distribution is normally distributed

```{r cars}


CombinedData[CombinedData==0]<-NA

summary(CombinedData$ABV)

hist(CombinedData$ABV)

qqnorm(CombinedData$ABV)

```

Question 7: Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatt10er plot.  Make your best judgment of a relationship and EXPLAIN your answer.

THe relationship between the bitterness and alcoholic content is positive. I am thinking the correlation is about 0.8.


```{r cars}

plot(CombinedData$ABV,CombinedData$IBU,pch = 15,xlab = "Alcoholic Content",ylab = "Bittternes", main = "Alcohol vs Bitternes")

```



