---
title: "Case Study 8 & 9"
output: html_document
---

```{r}
#Bring in Libraries
#install.packages('plyr')
#install.packages('tidyr')
#install.packages('dplyr')
#install.packages('tidyverse')
#install.packages('class')
#install.packages('caret')
#install.packages('e1071')
library(class)
library(caret)
library(e1071)
library(ggplot2)
library(plyr)     
library(tidyr)   
library(dplyr)
```

## R Markdown

Bring in Data

```{r}

#Pulled in the two csv documents
#Read.csv is the function i used to pull the documents
Beers_df=read.csv(file.choose(),header=TRUE)
Breweries_df=read.csv(file.choose(),header=TRUE)

```


Question 1: How many breweries are present in each state?

```{r}
#Created a dataframe  called Breweries_state to show the breweries in each State
Breweries_State <-aggregate(cbind(count = Name) ~ State,data = Breweries_df,FUN=function(x){NROW(x)})

#showed the top two rows
head(Breweries_df,2)
```



Divided the dataframe into two, so that we could fit all the states into one bar graph  

```{r}

# divided two dataframes into two

Breweries_df1<- head(Breweries_State,26)
Breweries_df2<-tail(Breweries_State,25)

```


The below shows a bar graph for the first half of the data

```{r}

ggplot(Breweries_df1, aes(x=reorder(State,-count), y=count))+geom_bar(stat = 'identity')+geom_text(aes(label=count),vjust=0, size=3.5)+ggtitle("State Vs Breweries")+xlab("State") + ylab("No Of Breweries")


```

The below shows a bar graph for the second half of the data

```{r}

ggplot(Breweries_df2, aes(x=reorder(State,-count), y=count))+geom_bar(stat = 'identity')+geom_text(aes(label=count),vjust=0, size=3.5)+ggtitle("State Vs Breweries")+xlab("State") + ylab("No Of Breweries")


```



Question 2   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)

```{r pressure, echo=FALSE}

#change the brew ID's header  from the Brewery_Table 
Breweries_df <- Breweries_df %>% rename(Brewery_id=Brew_ID)

#change the name column to Brewery_ID.Name to avoid duplicates columns
Breweries_df <- Breweries_df %>% rename(Brewery_id.Name=Name)

#Right join to the brewery data
CombinedData<-join(Beers_df,Breweries_df,by ="Brewery_id",type = "right")

#shows the top 2 
head(CombinedData,2)
```



Question 3: Address the missing values in each column.

```{r}

#Changed all the null values to NA
CombinedData[CombinedData=='']<-NA

#Count all the NA's
colSums(is.na(CombinedData))

```


Question 4: Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.


```{r}

#changed #NA to 0 to get actual medians
CombinedData[is.na(CombinedData)]<-0

#created the dataframes with medians for each State
Medians<-CombinedData%>%group_by(State)%>%summarise(median.ABV=median(ABV),median.IBU=median(IBU))

#top 3 data
head(Medians,3)
```


Divided the dataframes into two again to fit all the states in the bar graph 

```{r}

# divided two dataframes into two

Medians_df1<- head(Medians,26)
medians_df2<-tail(Medians,25)

```


The below shows a bar graph for ABV for the first half of the data

```{r}

ggplot(Medians_df1,aes(x=reorder(State,-median.ABV),y=median.ABV))+geom_bar(stat ='identity')+ggtitle("State vs ABV")+xlab("State")+ylab("Median Of ABV")


```



The below shows a bar graph for ABV for the Second half of the data

```{r}

ggplot(medians_df2,aes(x=reorder(State,-median.ABV),y=median.ABV))+geom_bar(stat ='identity')+ggtitle("State vs ABV")+xlab("State")+ylab("Median Of ABV")


```



The below shows a bar graph for IBU for the first half of the data

```{r}

ggplot(Medians_df1,aes(x=reorder(State,-median.IBU),y=median.IBU))+geom_bar(stat ='identity')+ggtitle("State vs IBU")+xlab("State")+ylab("Median Of IBU")


```


The below shows a bar graph for IBV for the Second half of the data

```{r}

ggplot(medians_df2,aes(x=reorder(State,-median.IBU),y=median.IBU))+geom_bar(stat ='identity')+ggtitle("State vs IBU")+xlab("State")+ylab("Median Of IBU")


```

Question 5: Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

the state with the highest ABV = Colorado
the state with the highest IBU = Oregon

```{r}

Maximum<-CombinedData%>%group_by(State)%>%summarise(Max.ABV=max(ABV),Max.IBU=max(IBU))

head(Maximum,5)
```


Question 6: Comment on the summary statistics and distribution of the ABV variable.

- ABV is normally in the range of 0.001 to 0.128. 
- But most of the beers in the country have a ABV of 0.059

- Based on the histogram and qqplot we can see that the distribution is normally distributed

```{r}


CombinedData[CombinedData==0]<-NA

summary(CombinedData$ABV)

```


Below is the histogram

```{r}

hist(CombinedData$ABV)

```

Below is the qqnorm

```{r}
qqnorm(CombinedData$ABV)

```



Question 7: Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatt10er plot.  Make your best judgment of a relationship and EXPLAIN your answer.

THe relationship between the bitterness and alcoholic content is positive. I am thinking the correlation is about 0.8.


```{r cars}

plot(CombinedData$ABV,CombinedData$IBU,pch = 15,xlab = "Alcoholic Content",ylab = "Bittternes", main = "Alcohol vs Bitternes")

```

8.  Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually.

In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged


```{r}

#first create the IPA and ALe Lables 

CombinedData$Label <- ifelse(grepl("IPA",CombinedData$Style, ignore.case=T),"IPA",ifelse(grepl("Ale",CombinedData$Style,ignore.case = T),"Ale","other"))


# filter data to IPA NA ALe

Model_Table <- CombinedData%>% filter(Label=="IPA"|Label=="Ale")

Model_Table<-Model_Table%>%drop_na()
```

IPA has stronger Alcohol Content

```{r}
Model_Table%>%ggplot(aes(x=Label,y=ABV,fill=Label))+geom_boxplot()

```



IPA's has a higher Bitter content
```{r}
Model_Table%>%ggplot(aes(x=Label,y=IBU,fill=Label))+geom_boxplot()

```
KNN model
```{r}

# create model


splitperc= .75
trainIndices = sample(1:dim(Model_Table)[1],round(splitperc*dim(Model_Table)[1]))
train = Model_Table[trainIndices,]
test = Model_Table[-trainIndices,]


# ran the model when k = 3

classifications = knn(train[,c(4,5)],test[,c(4,5)],train$Label, prob = TRUE, k = 3)
table(classifications,test$Label)
confusionMatrix(table(classifications,test$Label))
```


Ran the model when KNN = 5

```{r}
#k = 5

classifications = knn(train[,c(4,5)],test[,c(4,5)],train$Label, prob = TRUE, k =5)
table(classifications,test$Label)
confusionMatrix(table(classifications,test$Label))
```

We plotted a graph to show the best KNN


```{r}
accs = data.frame(accuracy = numeric(90), k = numeric(90))

for(i in 1:90)
{
  classifications = knn(train[,c(4,5)],test[,c(4,5)],train$Label, prob = TRUE, k = i)
  table(test$Label,classifications)
  CM = confusionMatrix(table(test$Label,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
plot(accs$k,accs$accuracy, type = "l", xlab = "k")


```
